---
title: 'Concentration of Attacks on Civilians in Darfur'
subtitle: '(2016 - 2021) - Visual 5'
author: "Rob Boswell"
date: "7-27-2023"
output: 
  html_document:
    css: rmd_subtitle_same_size_as_rmd_title.css
class_options: landscape
---

```{r echo=TRUE}

library(stringr)

# Get the current page number from the file name
current_page <- as.numeric(str_extract(knitr::current_input(), "\\d+"))

# Set the total number of pages
total_pages <- 9

# Generate the URLs for the previous and next pages
previous_page <- ifelse(current_page > 1, paste0("visual_", current_page - 1, "-darfur_violence-code_included.html"), NA)
next_page <- ifelse(current_page < total_pages, paste0("visual_", current_page + 1, "-darfur_violence-code_included.html"), NA)

```

<ul class="pager">
  <li><a href="`r previous_page`">Previous Page</a></li>
  <li><a href="`r next_page`">Next Page</a></li>
</ul>

<br>

```{r Maps: 2016-2021, echo=TRUE, fig.height=10.5, fig.width=10.5, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(ggplot2)
library(ggraph)
library(ggmap)
library(ggpubr)


setwd("C:/Users/rsb84/Desktop/RB/COLUMBIA/QMSS/COURSES/Spring_2021/Data Visualization/End_project")

ACLED_data <- readxl::read_excel("ACLED-DARFUR-VAC-2008-2021 (After Course Ended-for 2023 Portfolio)-UPDATED VERSION-inter1_numbers_replaced_with_actor_names.xlsx", 
     col_types = c("date", "numeric", "text", 
         "text", "text", "text", "text", "text", 
         "text", "text", "text", "text", "text", 
         "numeric", "numeric", "text", "text", 
         "text", "numeric", "numeric"))

acled_short=ACLED_data[c("event_date","year","sub_event_type","actor1","inter1","admin1","admin2","location","latitude","longitude","attacks","fatalities")]


# Below, I have created faceted heat density maps which allow for comparisons of the locations where violent attacks on civilians were most concentrated between 2016 and 2021 across Darfur, along with depictions of the locations of both operational and closed down UN bases

acled_2016 = subset(acled_short, year == 2016)
acled_2017 = subset(acled_short, year == 2017)
acled_2018 = subset(acled_short, year == 2018)
acled_2019 = subset(acled_short, year == 2019)
acled_2020 = subset(acled_short, year == 2020)
acled_2021 = subset(acled_short, year == 2021)


#Create columns that show which bases existed during which years

#2016:

geo_pko <- readxl::read_excel("active_bases-removed_bases_gone_more_than_6_months-Geo-PKO-v-2-1-2023_portfolio.xlsx")
geo_pko_for_maps_16 = filter(geo_pko, year == 2016)
geo_pko_df_16 = geo_pko_for_maps_16[c("year","location","latitude", "longitude")]
geo_pko_16_avg_troops_per_base = geo_pko_for_maps_16 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_16 = inner_join(geo_pko_16_avg_troops_per_base, geo_pko_df_16, by="location")
geo_pko_df_16 = distinct(geo_pko_df_16)
active_bases_2016 = geo_pko_df_16


#2017

geo_pko_for_maps_17 = filter(geo_pko, year == 2017)
geo_pko_df_17 = geo_pko_for_maps_17[c("year","location","latitude", "longitude")]
geo_pko_17_avg_troops_per_base = geo_pko_for_maps_17 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_17 = inner_join(geo_pko_17_avg_troops_per_base, geo_pko_df_17, by="location")
geo_pko_df_17 = distinct(geo_pko_df_17)
active_bases_2017 = geo_pko_df_17


#2018

geo_pko_for_maps_18 = filter(geo_pko, year == 2018)
geo_pko_df_18 = geo_pko_for_maps_18[c("year","location","latitude", "longitude")]
geo_pko_18_avg_troops_per_base = geo_pko_for_maps_18 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_18 = inner_join(geo_pko_18_avg_troops_per_base, geo_pko_df_18, by="location")
geo_pko_df_18 = distinct(geo_pko_df_18)
active_bases_2018 = geo_pko_df_18


#2019

geo_pko_for_maps_19 = filter(geo_pko, year == 2019)
geo_pko_df_19 = geo_pko_for_maps_19[c("year","location","latitude", "longitude")]
geo_pko_19_avg_troops_per_base = geo_pko_for_maps_19 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_19 = inner_join(geo_pko_19_avg_troops_per_base, geo_pko_df_19, by="location")
geo_pko_df_19 = distinct(geo_pko_df_19)
active_bases_2019 = geo_pko_df_19


#2020

geo_pko_for_maps_20 = filter(geo_pko, year == 2020)
geo_pko_df_20 = geo_pko_for_maps_20[c("year","location","latitude", "longitude")]
geo_pko_20_avg_troops_per_base = geo_pko_for_maps_20 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_20 = inner_join(geo_pko_20_avg_troops_per_base, geo_pko_df_20, by="location")
geo_pko_df_20 = distinct(geo_pko_df_20)
active_bases_2020 = geo_pko_df_20


#2021

geo_pko_for_maps_21 = filter(geo_pko, year == 2021)
geo_pko_df_21 = geo_pko_for_maps_21[c("year","location","latitude", "longitude")]
geo_pko_21_avg_troops_per_base = geo_pko_for_maps_21 %>% group_by(location) %>% summarise(avg.troops = mean(no.troops, na.rm=T))
geo_pko_df_21 = inner_join(geo_pko_21_avg_troops_per_base, geo_pko_df_21, by="location")
geo_pko_df_21 = distinct(geo_pko_df_21)
active_bases_2021 = geo_pko_df_21



#Note: In my 2023 portfolio version of the following closed UN bases files, I do not include sites that were closed in July or after of the given year since July is the 7th month of the year - meaning such bases were active for at least half of the given year.


#Sources for the dates of UN base closures:
#https://oios.un.org/file/8866/download?token=F6tc9ZI5
#https://www.ecoi.net/en/file/local/2052225/S_2021_470_E.pdf
#https://peacekeeping.un.org/en/unamid-hands-back-former-iss-compound-to-agricultural-research-centre-el-fasher-north-darfur
#https://peacekeeping.un.org/en/unamid-expresses-concern-over-allegations-of-improper-team-site-handovers


closed_bases_2017 <- read_csv("closed_bases_2017-2023_portfolio.csv")

closed_bases_2018 <- read_csv("closed_bases_2018-2023_portfolio.csv")

closed_bases_2019 <- read_csv("closed_bases_2019-2023_portfolio.csv")

closed_bases_2020 <- read_csv("closed_bases_2020-2023_portfolio.csv")

closed_bases_2021 <- read_csv("closed_bases_2021-2023_portfolio.csv")


#MAPS FROM 2016-2021, FACETED

map_darfur <- get_map(location=c(21.71368, 8.64132, 27.58313, 16.12054), zoom=7, source = "stamen", maptype="terrain")


map_2016 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur",        fontface=2,size=3) + 
annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2016, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2016, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=22, color="steelblue") +
guides(size="none")  + 
theme(legend.position="bottom", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2016", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_2017 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur", fontface=2,size=3) + annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2017, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2017, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=22, color="steelblue") + 
geom_point(data = closed_bases_2018, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=25, color="darkgreen") +
guides(size="none") +
guides(size="none")  + 
theme(legend.position="none", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2017", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_2018 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur",        fontface=2,size=3) + 
annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2018, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2018, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.5), shape=22, color="steelblue") +
guides(size="none")  + 
geom_point(data = closed_bases_2018, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=25, color="darkgreen") +
guides(size="none")  +
theme(legend.position="none", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2018", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_2019 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur",        fontface=2,size=3) + 
annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2019, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2019, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.5), shape=22, color="steelblue") +
guides(size="none")  + 
geom_point(data = closed_bases_2019, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=25, color="darkgreen") +
guides(size="none")  +
theme(legend.position="none", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2019", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_2020 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur",        fontface=2,size=3) + 
annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2020, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2020, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.5), shape=22, color="steelblue") +
guides(size="none")  + 
geom_point(data = closed_bases_2020, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=25, color="darkgreen") +
guides(size="none")  +
theme(legend.position="none", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2020", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_2021 = ggmap(map_darfur) + 
plot.new()+ 
scale_fill_gradient(low="yellow", high="red") + 
annotate("text", x= 26.609638, y= 10.817195, label="East Darfur", fontface=2,size=3) + 
annotate("text", x= 25.556517, y= 16.217571, label="North Darfur", fontface=2,size=3) + 
annotate("text", x= 24.420469, y= 11.002820, label="South Darfur", fontface=2,size=3) + 
annotate("text", x= 23.116657, y= 14.519303, label="West Darfur", fontface=2,size=3) + 
stat_density2d(data=acled_2021, geom = "polygon", aes(x=longitude, y=latitude,fill=..level.., alpha=..level..)) + 
guides(alpha="none") + 
geom_point(data = active_bases_2021, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.5), shape=22, color="steelblue") +
guides(size="none")  + 
geom_point(data = closed_bases_2021, mapping = aes(x = longitude, y = latitude, size=1.5, alpha=0.3), shape=25, color="darkgreen") +
guides(size="none")  +
theme(legend.position="none", title =element_text(size=13, color=  "steelblue",face='bold'), axis.title.x=element_blank(),              axis.title.y=element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), rect = element_blank()) + 
labs(title="2021", subtitle="", caption = "") + 
guides(fill=guide_legend(title="Concentration of All\nAttacks on Civilians"))


map_facet <- ggarrange(map_2016, map_2017, map_2018, map_2019, map_2020, map_2021, heights = c(.8, .8, .8, .8, .8, .8), widths = c(.8,.8,.8,.8,.8,.8), align = "h", nrow = 2, ncol=3, common.legend = TRUE, legend="bottom")

```

<center>
```{r Faceted Maps: 2016-2021, echo=TRUE, fig.height=10.5, fig.width=10.5, message=FALSE, warning=FALSE}

annotate_figure(map_facet, top = text_grob("Concentration of All Attacks on Darfur Civilians, 2016-2021", face = "bold", size = 26, color="steelblue", vjust = 0.5), bottom = text_grob("Note: Squares are Active UN Bases. Triangles are Closed Down Bases", size = 12, color ="steelblue"))

```
</center>

<br>

#### **Takeaways:** It can be seen that the density of attacks on civilians prior to the start of closures of UNAMID bases in 2017, and up until 2018, was primarily in the region where the borders of Central, North, and South Darfur meet. However, the concentration of attacks substantially condensed into the northern part of South Darfur in 2019 - the year in which the UN formally decided to withdraw the mission - before suddenly expanding again in 2020 back into Central, North, and West Darfur. The expansion into West Darfur in 2020 appears particularly noteable since there had not been much of a concentration of attacks in West Darfur until that year. 

<br>

#### It is possible that this expansion may have been partly a result of the UN closing down 8 of its bases in West Darfur from 2017-2019. Potential causes of this alteration in the density deserve more inspection. Back in Visual 3, we saw that an unusually large spike in civilian killings in Western Darfur occured in 2019, and that by 2021 there were more civilian killings there (155) than in any other Darfur region. Although the 2019 and 2021 density maps here do not show the highest concentration of attacks on civilians being in West Darfur, what likely accounts for the discrepancy is that there may have been an overall large number of fatalities, on average, per attack.

<br>

#### In the next visual (Visual 6), I have created a faceted comparison map of Darfur between the years 2016 and 2020, with bubble size representing the number of fatalities per attack on civilians, and with varying shades of green coloring in administrative regions reflecting the average number of UN troops present, in order to discern overall trends.

```{r echo=TRUE}

# Get the current page number from the file name
current_page <- as.numeric(str_extract(knitr::current_input(), "\\d+"))

# Set the total number of pages
total_pages <- 9

# Generate the URLs for the previous and next pages
previous_page <- ifelse(current_page > 1, paste0("visual_", current_page - 1, "-darfur_violence-code_included.html"), NA)
next_page <- ifelse(current_page < total_pages, paste0("visual_", current_page + 1, "-darfur_violence-code_included.html"), NA)

```

<ul class="pager">
  <li><a href="`r previous_page`">Previous Page</a></li>
  <li><a href="`r next_page`">Next Page</a></li>
</ul>
